Connectome inferences using partial correlation from imaging calcium signal

### 2014-06-13 article.tex
wording

\subsection{Causal discovery methodology}
% Introduce a custom solution to make the Y matrix asymmetric
% Introduce directivity
Since the partial correlation is a symmetric measure, the causal mechanism behind the
interaction ($A$ directly causes $B$, $B$ directly causes $A$ or both) can not
be directly inferred\footnote{Also note that two other causal mechanisms might be
implied while having a non-zero partial coefficient: (i) a pair of variables
is induced by a common (hidden) variable; (ii) $A$ (respectively $B$) is
conditionally correlated to a hidden variable affecting $B$ (resp. $A$)
\cite{de2004discovery}.} from a symmetric adjacency matrix.

By ignoring the directivity, we systematically make a true positive and a
false positive for direct link (see a more detailed discussion about the
chosen metric in section \ref{sec:metric}). Therefore, performance might be
improved by predicting an asymmetric adjacency matrix. Hoping to orientate
correctly the most obvious directed links, we attempt to retrieve some causal
information, i.e. the directivity of the links, by computing and stacking with
our symmetric adjacency matrix a matrix of pairwise activation heuristics.
This method computes an anti-symmetric activation score $z_{ij}$ based the
variation of fluorescence signal of a neuron $j$ due to a neuron $i$ over all
timesteps:
\[
z_{ij} = \sum_{t=1}^n \mathbb{1}(X_{t+1,j} \ge X_{t, i}) \cdot
\mathbb{1}(f_1\le X_t^i \le f_2) -  \mathbb{1}(X_{t+1,i} \ge X_{t, j}) \cdot
\mathbb{1}(f_1 \le X_t^j \le f_2)
\]
where $f_1$ and $f_2$ are parameters of the method and $I$ is the indicator
function. Prior computing $z_{ij}$ only the filter defined by Equations
\ref{eq:low-pass}, \ref{eq:high-pass} and \ref{eq:hard-treshold-filter} were
used.

@@

% Metrics : ROC AUC and AUPRC
In supervised learning terms, the network inference task can be viewed as a
binary classification task where one has to correctly classify the presence
or the absence of edges of the graph. Given the ground true adjacency matrix
$A$ and a connectivity score matrix $\hat{A}$, we assess the accuracy of the
network inference method using the area under the ROC curve (ROC AUC)
and the area under the precision-recall curve (AUPRC)
(\cite{schrynemackers2013protocols}). Since neurons are not self-connected
($(i, i) \not \in E, \forall i \in V$) and some inference methods could
give maximal scores to the diagonal of $\hat{A}$, e.g. Pearson
correlation, we set those elements to the minimum of $\hat{A}$ prior evaluating
those algorithms.

>>>>

\subsection{Causal discovery methodology}
% Introduce a custom solution to make the Y matrix asymmetric
% Introduce directivity
Since partial correlation is a symmetric measure, the causal mechanism behind the
interaction ($A$ directly causes $B$, $B$ directly causes $A$ or both) can not
be directly inferred\footnote{Also note that two other causal mechanisms might be
implied while having a non-zero partial coefficient: (i) a pair of variables
is induced by a common (hidden) variable; (ii) $A$ (respectively $B$) is
conditionally correlated to a hidden variable affecting $B$ (resp. $A$)
\cite{de2004discovery}.}.

By ignoring the directivity, we systematically make a true positive and a
false positive for direct link.  Hoping to orientate
correctly the most obvious directed links, we attempt to retrieve some causal
information, i.e. the directivity of the links, by computing and stacking with
our symmetric adjacency matrix a matrix of pairwise activation heuristics.
This method computes an anti-symmetric activation score $z_{ij}$ based the
variation of fluorescence signal of a neuron $j$ due to a neuron $i$ over all
timesteps:
\[
z_{ij} = \sum_{t=1}^n \mathbb{1}(X_{t+1,j} \ge X_{t, i}) \cdot
\mathbb{1}(f_1\le X_t^i \le f_2) -  \mathbb{1}(X_{t+1,i} \ge X_{t, j}) \cdot
\mathbb{1}(f_1 \le X_t^j \le f_2)
\]
where $f_1$ and $f_2$ are parameters of the method and $I$ is the indicator
function. Prior computing $z_{ij}$ only the filter defined by Equations
\ref{eq:low-pass}, \ref{eq:high-pass} and \ref{eq:hard-treshold-filter} were
used.

@@

% Metrics : ROC AUC and AUPRC
In supervised learning terms, the network inference task can be viewed as a
binary classification task where one has to correctly classify the presence
or the absence of edges of the graph. Given the ground true adjacency matrix
$A$ and a connectivity score matrix $\hat{A}$, we assess the accuracy of the
network inference method using the area under the ROC curve (ROC AUC)
and the area under the precision-recall curve (AUPRC)
(\cite{schrynemackers2013protocols}). Since neurons are not self-connected
($(i, i) \not \in E, \forall i \in V$) and some inference methods could
give maximal scores to the diagonal of $\hat{A}$, e.g. Pearson
correlation, we set those elements to the minimum of $\hat{A}$ prior
assessment.
